---
title: "Guide to `phipmake` for Larman Lab"
author: "Brandon Sie"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This guide is intended to help members of the [Larman Lab](http://labs.pathology.jhu.edu/larman/) at JHU to get up and running with `phipmake` on MARCC, with help of the SLURM task scheudler. Some of the steps are generalizable and others are specific to the Larman Lab.

Some definitions:  

* [`drake`](https://github.com/ropensci/drake) is an R package written by [Will Landau](https://github.com/wlandau) and part of [rOpenSci](https://github.com/ropensci) designed to support reproducible data analysis workflows. `drake` keep track of which steps of a data analysis pipeline are out-of-date, so that you can update only those steps without re-running all of your code from scratch.  
* [`phipmake`](https://github.com/brandonsie/phipmake) is an R package written to use tools from `drake` to implement the Larman Lab's PhIP-Seq analysis pipeline.
* _plans_ are workflow recipes, consisting of objects to be build (targets) and instructions for building those objects (commands).  
* _targets_ are individual objects in a `drake` workflow to be built.  

<!-- counts  -->
<!-- enrichments -->
<!-- prosum -->
<!-- promax -->
<!-- polyclonal -->

<!-- ## Prerequisites -->
<!-- marcc account -->
<!-- blast+ -->
<!-- library annotations etc -->

# Setup


## 1. MARCC  

For usage with PhIPdb, the following R package installation steps should occur via a secure shell (SSH) connection to MARCC, using an account that is part of the hlarman1 group. For each step, I provide an option using interactive mode and using MARCC's SLURM task scheduler. 

To install packages interactively without using the login node, use an `srun` command similar to the following:
```{command1, eval = FALSE}
srun -N 1 -p shared --pty bash
ml R
R
```

To install packages using the SLURM scheduler, use the resources in PhIPdb/Software/install_r_pkg:
```{command2, eval = FALSE}
# Navigate to install_r_pkg
cd /data/hlarman1/PhIPdb/Software/install_r_pkg

# Example of installing a package hosted on CRAN
sbatch --export=pkg="tidyverse" installCRAN.sh

# Example of installing a package hosted on GitHub
sbatch --export=pkg="brandonsie/phipmake" installGithub.sh

```


## 2. Install [`drake`](https://github.com/ropensci/drake) and [`phipmake`](https://github.com/brandonsie/phipmake) from GitHub:  

```{r, eval = FALSE}
# Interactive
if(!requireNamespace("remotes")) install.packages("remotes")

# drake is on CRAN, but phipmake depends on the more recently updated Github version.
remotes::install_github("ropensci/drake") 
remotes::install_github("brandonsie/phipmake")

```

```{command3, eval = FALSE}
# SLURM
sbatch --export=pkg="ropensci/drake" installGithub.sh
sbatch --export=pkg="brandonsie/phipmake" installGithub.sh

```


This should install `drake`, `phipmake`, and all required dependencies.

## 3. Install optional depdendencies for parallelization.  

```{r, eval = FALSE}
install.packages("future")
install.packages("future.apply")

```

```{command4, eval = FALSE}
# SLURM
sbatch --export=pkg="future" installCRAN.sh
sbatch --export=pkg="future.apply" installCRAN.sh

```


`drake` uses `future` to parellalize plans so that multiple targets can be generated simultaneously. `phipmake` uses `future.apply` to parallelize within certain individual targets.

# Usage on MARCC

## Parameter Setup  

`phipmake` relies on a parameters file, usually a file called drake_params.tsv and stored in the root directory of a project in PhIPdb/ProcessedData. For example, PhIPdb/ProcessedData/phipseq_0100/drake_params.tsv.

To create drake_params.tsv with default parameters for a new PhIP-Seq screen interactively, you can use the following command from `phipmake`. Alternatively, it's fine to copy and modify a drake_params.tsv file from a previous project.:
```{r, eval = FALSE}
# Replace phipseq_9999 with appropriate project ID.
phipmake:::write_drake_params(
  dir = "/data/hlarman1/PhIPdb/ProcessedData/phipseq_9999",
  screen_name = "phipseq_9999")
```

This will write drake_params.tsv to the specified directory, expecting a counts file named counts.csv, an enrichment file called enrichment.csv. `write_drake_params` also assumes a default enrichment threshold of 5, a metadata path of /data/hlarman1/PhIPdb/Metadata/PeptideLibraries, and a desired output file extension of tsv. All of these options can be modified via parameters passed to `write_drake_params.`

## Running `phipmake`

Navigate to the phipmake software directory in PhIPdb and use the provided shell script.
```{command5, eval = FALSE}
cd /data/hlarman1/PhIPdb/Software/phipmake


#Setup variables for sbatch exports
pdir="/home-1/bsie1@jhu.edu/data/PhIPdb/ProcessedData/"
screen="phipseq_0101" 
plans="Counts-Enrichment-Polyclonal-AVARDA" 
targs_to_clean="NULL" 
pjobs=4 

# Some notes about above variables
# screen: changes from project to project
# plans and targs_to_clean four possible plans are delimited by "-". See additional notes below.
# pjobs: can be set to 1 to not parallelize and not depeend on future or future.apply

sbatch --export=wd=$pdir$screen,plan=$plans,clean=$targs_to_clean,njobs=$pjobs runphipmake.sh
```

Explanation of some of the parameters exported in the `sbatch` command:  

* `plans` specifies the four sub-workflows in the Larman Lab's `phipmake` pipeline that can be toggled on/off. Targets are delimited by hyphens "-". For example, to run only Enrichment and Polyclonal parts of the pipeline, excluding Counts and AVARDA, set `plans="Enrichment-Polyclonal"`.   
* `targs_to_clean` specifies targets of the `phipmake` plan to force to be rebuilt, even if `drake` thinks that that target is up-to-date. These targets are also delimited by hyphens "-". For example, to indicate that data pulled from annotation files must be updated, set `targs_to_clean="counts_annotations-enrichment_annotations"`. This will delete drake's cache for these two targets, and so these targets and all downstream dependencies should be rebuilt.


# Types of Output Files

# Maintenance & Updating

## Updating annotation files


## Adding a New Library

### BLAST+, Tidying
 
### Annotations

## Github Pull Requests  

PR's are welcome.